\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-03-29 paper.qmd
Added latest citation for non-metric MDS
Initial notation table set up

<!-- - Connection between NLDR and MDS-->
Historically, $k-D$ representations of $p-D$ data have been computed using multidimensional scaling (MDS) [@borg2005], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.  The $k-D$ representation can be considered to be a layout of points in $k-D$ produced by an embedding procedure that maps the data from $p-D$. In MDS, the $k-D$ layout is constructed by minimizing a stress function that differences distances between points in $p-D$ with potential distances between points in $k-D$. Various formulations of the stress function result in non-metric scaling [@kruskal1964] and isomap [@silva2002]. Challenges in working with high-dimensional data, including visualization, are outlined in @johnstone2009.

@@

>>>>

<!-- - Connection between NLDR and MDS-->
Historically, $k-D$ representations of $p-D$ data have been computed using multidimensional scaling (MDS) [@borg2005], which includes principal components analysis (PCA) [@jolliffe2011] as a special case.  The $k-D$ representation can be considered to be a layout of points in $k-D$ produced by an embedding procedure that maps the data from $p-D$. In MDS, the $k-D$ layout is constructed by minimizing a stress function that differences distances between points in $p-D$ with potential distances between points in $k-D$. Various formulations of the stress function result in non-metric scaling [@saeed2018] and isomap [@silva2002]. Challenges in working with high-dimensional data, including visualization, are outlined in @johnstone2009.

@@

```{r}
# Notations used in the paper

notation_df <- read_csv("data/paper_notations.csv")

# Create the table
kable(notation_df, format = "latex", booktabs = TRUE, escape = FALSE, 
      caption = "Notations used in this paper.", label = "tbl-notations") |>
  kable_styling(full_width = FALSE) |>
  row_spec(0, bold = TRUE)
```

### 2024-03-29 paper.tex
Corrected ref error
Initial notation table set up

Historically, \(k-D\) representations of \(p-D\) data have been computed
using multidimensional scaling (MDS) \citep{borg2005}, which includes
principal components analysis (PCA) \citep{jolliffe2011} as a special
case. The \(k-D\) representation can be considered to be a layout of
points in \(k-D\) produced by an embedding procedure that maps the data
from \(p-D\). In MDS, the \(k-D\) layout is constructed by minimizing a
stress function that differences distances between points in \(p-D\)
with potential distances between points in \(k-D\). Various formulations
of the stress function result in non-metric scaling \citep{kruskal1964}
and isomap \citep{silva2002}. Challenges in working with
high-dimensional data, including visualization, are outlined in
\citet{johnstone2009}.

@@

>>>>

Historically, \(k-D\) representations of \(p-D\) data have been computed
using multidimensional scaling (MDS) \citep{borg2005}, which includes
principal components analysis (PCA) \citep{jolliffe2011} as a special
case. The \(k-D\) representation can be considered to be a layout of
points in \(k-D\) produced by an embedding procedure that maps the data
from \(p-D\). In MDS, the \(k-D\) layout is constructed by minimizing a
stress function that differences distances between points in \(p-D\)
with potential distances between points in \(k-D\). Various formulations
of the stress function result in non-metric scaling \citep{saeed2018}
and isomap \citep{silva2002}. Challenges in working with
high-dimensional data, including visualization, are outlined in
\citet{johnstone2009}.

@@

\begin{table}

\caption{\label{tab:tbl-notations}Notations used in this paper.}
\centering
\begin{tabular}[t]{ll}
\toprule
\textbf{Notation} & \textbf{Description}\\
\midrule
$n$ & number of observations\\
$p$ & number of dimensions in high-D data\\
$X_{n \times p} = \begin{bmatrix} \textbf{x} _{1} & \textbf{x}_ {2} & \cdots & \textbf{x}_{n} \\  \end{bmatrix}^\top$ & high-D data matrix\\
$d$ & number of dimensions in embedding data, usually two\\
$Y_{n \times d} = \begin{bmatrix} \textbf{y} _{1} & \textbf{y}_ {2} & \cdots & \textbf{y}_{n} \\  \end{bmatrix}^\top$ & embedding matrix\\
\addlinespace
$P_t$ & point along the geodesic path at time $t$\\
$P_{t+1}$ & point along the geodesic path at the next time instant, which is $t+1$\\
$b_1$ & number of bins along the x-axis\\
$b_2$ & number of bins along the y-axis\\
$b$ & total number of bins in the hexagonal grid\\
\addlinespace
$b'$ & total number of bins without the empty bins\\
$C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})$ & hexagonal bin centroid coordinates\\
$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ & averaged high-D coordinates\\
\bottomrule
\end{tabular}
\end{table}
